---
title: "Gene Set Enrichment Analysis"
date: "May 07, 2019"
output: html_document
---

```{r echo=F,message=F,warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


# Changes from previous version


1) There is now a column `method`, indicating whether a linear model or logistic regression was used
2) For genes with multiple SNPs, the median value of `prob_Z1_conf` was used.


# Birthweight T2D


```{r,message=F,warning=F}
library(tidyverse)
library(drake)


gene_summ_df <- readd("summ_cause_med_cause_df_causedata_1L")
gr <- readd("genres_dfmed_1L_summ_cause_med_cause_df_causedata_1L_medz_causedata_1L")
gene_df <- readd("genresmed_summ_cause_med_cause_df_causedata_1L_genres_dfmed_1L_summ_cause_med_cause_df_causedata_1L_medz_causedata_1L")
shuff <- readd("genres_dfmedshuff_1L_summ_cause_med_cause_df_causedata_1L_medz_causedata_1L")
loadd(godf)

go_names <- distinct(godf,go_id,name_1006) %>% rename(GO_term=name_1006)
```



```{r}
dplyr::filter(gr,term!="(Intercept)",p_BH<0.05) %>% 
  dplyr::select(-file) %>%  
  inner_join(go_names,by="go_id") %>% 
  dplyr::select(-term) %>% 
  arrange(p_BH) %>% 
  DT::datatable(caption="Enriched Pathways(BW->T2D)")
```

### genes in enriched pathways 


```{r}
gene_df %>% DT::datatable(caption="Genes in Enriched Pathways(BW->T2D)")
```


## Permuting cause results


For the birthweight T2D results I permuted the `prob_Z1_conf` vector 100 times and repeated the analysis. Below I plot the original asymptotic $p$  and the permutation $p$ (i.e the proportion of permuted $p$ values less than the non-permuted $p$).  With the exception of two logistic $p$-values, they are remarkably consistent. 

```{r permutationgen}
perm_df <- map_df(1:100L,function(x){
  readd(glue::glue("genres_dfmedshuff_{x}L"),character_only = TRUE)
})

my_df <- rename_at(gr,vars(-term,-method,-go_id,-file),function(x){paste0(x,"_0")}) %>%
  inner_join(perm_df)

p_res <- filter(my_df,term!="(Intercept)",go_id!="",!is.na(p.value_0)) %>%
  group_by(method,go_id) %>% 
  summarize(
    p_p=mean(p.value<p.value_0,na.rm=T)) %>% 
  ungroup()

new_gr <- inner_join(gr,p_res) %>% rename(p_permutation=p_p)

filter(new_gr,term!="(Intercept)") %>% ggplot(aes(x=p_permutation,y=p.value))+
  geom_point()+xlab("Permutation P")+ylab("Asymptotic P")+
  facet_wrap(~method)+
  geom_abline(slope=1,intercept=0)

```

```{r}
filter(new_gr,term!="(Intercept)") %>% ggplot(aes(x=p_permutation,y=p.value))+
  geom_point()+xlab("Permutation P")+ylab("Asymptotic P")+
  facet_wrap(~method)+
  geom_abline(slope=1,intercept=0)+scale_x_log10()+scale_y_log10()
```



```{r}
dplyr::filter(new_gr,term!="(Intercept)",p_BH<0.05) %>% 
  dplyr::select(-file) %>%  
  inner_join(go_names,by="go_id") %>% 
  dplyr::select(-term) %>% 
  arrange(p_BH) %>% 
  DT::datatable(caption="(Permuted) Enriched Pathways(Height->CAD)")
```


# Height CAD




```{r,message=F,warning=F}
library(tidyverse)
library(drake)


gene_summ_df <- readd("summ_cause_med_cause_df_causedata_2L")
gr <- readd("genres_dfmed_2L_summ_cause_med_cause_df_causedata_2L_medz_causedata_2L")
gene_df <- readd("genresmed_summ_cause_med_cause_df_causedata_2L_genres_dfmed_2L_summ_cause_med_cause_df_causedata_2L_medz_causedata_2L")



stopifnot(gr$file[1]=="giant_height__vanderHarst_cad_cause.RDS")
```


### Enriched pathways
```{r}
dplyr::filter(gr,term!="(Intercept)",p_BH<0.05) %>% 
  dplyr::select(-file) %>%  
  inner_join(go_names,by="go_id") %>% 
  dplyr::select(-term) %>% 
  arrange(p_BH) %>% 
  DT::datatable(caption="Enriched Pathways(Height->CAD)")
```

### genes in enriched pathways 


```{r}
gene_df %>% DT::datatable(caption="Genes in Enriched Pathways(Height->CAD)")
```




# HDL CAD

```{r,message=F,warning=F}
library(tidyverse)
library(drake)


gene_summ_df <- readd("summ_cause_med_cause_df_causedata_3L")
gr <- readd("genres_dfmed_3L_summ_cause_med_cause_df_causedata_3L_medz_causedata_3L")
gene_df <- readd("genresmed_summ_cause_med_cause_df_causedata_3L_genres_dfmed_3L_summ_cause_med_cause_df_causedata_3L_medz_causedata_3L")



stopifnot(gr$file[1]=="glg_hdl__vanderHarst_cad_cause.RDS")

```



### Enriched pathways
```{r}
dplyr::filter(gr,term!="(Intercept)",p_BH<0.05) %>% 
  dplyr::select(-file) %>%  
  inner_join(go_names,by="go_id") %>% 
  dplyr::select(-term) %>% 
  arrange(p_BH) %>% 
  DT::datatable(caption="Enriched Pathways(HDL->CAD)")
```

### genes in enriched pathways 


```{r}
gene_df %>% DT::datatable(caption="Genes in Enriched Pathways(HDL->CAD)")
```



# LDL T2D

```{r,message=F,warning=F}
library(tidyverse)
library(drake)


gene_summ_df <- readd("summ_cause_med_cause_df_causedata_4L")
gr <- readd("genres_dfmed_4L_summ_cause_med_cause_df_causedata_4L_medz_causedata_4L")
gene_df <- readd("genresmed_summ_cause_med_cause_df_causedata_4L_genres_dfmed_4L_summ_cause_med_cause_df_causedata_4L_medz_causedata_4L")


stopifnot(gr$file[1]=="glg_ldl__diagram_t2d_cause.RDS")


```



### Enriched pathways
```{r ldlt2dp}
dplyr::filter(gr,term!="(Intercept)",p_BH<0.05) %>% 
  dplyr::select(-file) %>%  
  inner_join(go_names,by="go_id") %>% 
  dplyr::select(-term) %>% 
  arrange(p_BH) %>% 
  DT::datatable(caption="Enriched Pathways(LDL->T2D)")
  
```

### genes in enriched pathways 


```{r ldl_t2dg}
gene_df %>% DT::datatable(caption="Genes in Enriched Pathways(LDL->T2D)")
```





