---
title: "Analysis of 16 GWAS traits"
author: "Jean Morrison"
date: "May 18, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

We analyzed 16 of the GWAS traits analyzed by Pickerell et al. For each pair of traits, I fit four models using the two factor model:

1. $\lambda_{12} = \lambda_{21} = 0$
2. $\lambda_{12} = 0$, $\lambda_{21}$ unconstrained (allow an effect form trait 1 to trait 2)
3. $\lambda_{21} = 0$, $\lambda_{12}$ unconstrained (allow an effect form trait 2 to trait 1)
4. $\lambda_{12}, \lambda_{21}$ both unconstrained (allow effects in both directions)

For these fits I set the distribution of $\alpha_1$ and $\alpha_2$ using the ASH priors estimated for traits 1 and traits 2 separately. I assume the priors are independent.

## Results on pruned data set 

Effect sizes; data pruned for LD at a theshold of 0.1 and with LDAK weights:
```{r, echo=FALSE, show="hold"}
library(ggplot2)
library(reshape2)
library(gridExtra)
res <- readRDS("../data/all_res.prune_wts.RData")
resJ <- readRDS("../data/all_res.prune_wts.joint.RData")

#indep bic
pvals <- 2*pnorm(abs(res[1, , ]), lower.tail=FALSE)
logpvals <- -1*log10(pvals)

logpvals[(res[1, , ] < 0 & !is.na(res[1, , ])) ] <- -1 * logpvals[(!is.na(res[1, ,]) & res[1, ,] < 0)]

 dat.m <- melt(t(logpvals))
 dat.m$value[dat.m$value > 10] <- 10
 dat.m$value[dat.m$value < -10] <- -10
 dat.m$value[abs(dat.m$value) < 2] <- 0
 (pbic <- ggplot(dat.m, aes(x=Var1, y=Var2, fill=value)) + 
   geom_tile() + coord_equal() + ggtitle("Pruned, Best model(BIC)\nIndependent Prior") + 
   scale_fill_gradient2(low="blue", high="red") + 
   theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none"))
#joint bic
 pvals <- 2*pnorm(abs(resJ[1, , ]), lower.tail=FALSE)
logpvals <- -1*log10(pvals)

logpvals[(resJ[1, , ] < 0 & !is.na(resJ[1, , ])) ] <- -1 * logpvals[(!is.na(resJ[1, ,]) & resJ[1, ,] < 0)]

 dat.m <- melt(t(logpvals))
 dat.m$value[dat.m$value > 10] <- 10
 dat.m$value[dat.m$value < -10] <- -10
 dat.m$value[abs(dat.m$value) < 2] <- 0
 (pbic <- ggplot(dat.m, aes(x=Var1, y=Var2, fill=value)) + 
   geom_tile() + coord_equal() + ggtitle("Pruned, Best model(BIC)\nJoint Prior") + 
   scale_fill_gradient2(low="blue", high="red") + 
   theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none"))
 
```


```{r, echo=FALSE, eval=FALSE}
 #LRT indep
 pvals <- 2*pnorm(abs(res[2, , ]), lower.tail=FALSE)
logpvals <- -1*log10(pvals)

logpvals[(res[2, , ] < 0 & !is.na(res[2, , ])) ] <- -1 * logpvals[(!is.na(res[2, ,]) & res[2, ,] < 0)]

 dat.m <- melt(t(logpvals))
 dat.m$value[dat.m$value > 10] <- 10
 dat.m$value[dat.m$value < -10] <- -10
 dat.m$value[abs(dat.m$value) < 2] <- 0
 (plrt <- ggplot(dat.m, aes(x=Var1, y=Var2, fill=value)) + 
   geom_tile() + coord_equal() + ggtitle("Pruned, Best model(LRT)\nIndependent Prior") + 
   scale_fill_gradient2(low="blue", high="red") + 
   theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none"))
 
  #LRT joint
 pvals <- 2*pnorm(abs(resJ[2, , ]), lower.tail=FALSE)
logpvals <- -1*log10(pvals)

logpvals[(resJ[2, , ] < 0 & !is.na(resJ[2, , ])) ] <- -1 * logpvals[(!is.na(resJ[2, ,]) & resJ[2, ,] < 0)]

 dat.m <- melt(t(logpvals))
 dat.m$value[dat.m$value > 10] <- 10
 dat.m$value[dat.m$value < -10] <- -10
 dat.m$value[abs(dat.m$value) < 2] <- 0
 (plrt <- ggplot(dat.m, aes(x=Var1, y=Var2, fill=value)) + 
   geom_tile() + coord_equal() + ggtitle("Pruned, Best model(LRT)\nJoint Prior") + 
   scale_fill_gradient2(low="blue", high="red") + 
   theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none"))
 
 
```
 
 
 
```{r, echo=FALSE}
 pvals <- 2*pnorm(abs(res[3, , ]), lower.tail=FALSE)
logpvals <- -1*log10(pvals)

logpvals[(res[3, , ] < 0 & !is.na(res[3, , ])) ] <- -1 * logpvals[(!is.na(res[3, ,]) & res[3, ,] < 0)]

 dat.m <- melt(t(logpvals))
 dat.m$value[dat.m$value > 10] <- 10
 dat.m$value[dat.m$value < -10] <- -10
 dat.m$value[abs(dat.m$value) < 2] <- 0
 (pfull <- ggplot(dat.m, aes(x=Var1, y=Var2, fill=value)) + 
   geom_tile() + coord_equal() + ggtitle("Pruned, Full model\nIndependent Prior") + 
   scale_fill_gradient2(low="blue", high="red") + 
   theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none"))
 
 
 
  pvals <- 2*pnorm(abs(resJ[3, , ]), lower.tail=FALSE)
logpvals <- -1*log10(pvals)

logpvals[(resJ[3, , ] < 0 & !is.na(resJ[3, , ])) ] <- -1 * logpvals[(!is.na(resJ[3, ,]) & resJ[3, ,] < 0)]

 dat.m <- melt(t(logpvals))
 dat.m$value[dat.m$value > 10] <- 10
 dat.m$value[dat.m$value < -10] <- -10
 dat.m$value[abs(dat.m$value) < 2] <- 0
 (pfull <- ggplot(dat.m, aes(x=Var1, y=Var2, fill=value)) + 
   geom_tile() + coord_equal() + ggtitle("Pruned, Full model\nJoint Prior") + 
   scale_fill_gradient2(low="blue", high="red") + 
   theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none"))
 
 

```




A handmade list of differences from Pickerell et al:

Trait 1 | Trait 2 | Best BIC (Indep Prior) |  Best BIC (Joint Prior) | Pickerell et al
--------|---------|------------------------|-------------------------|-----------------
scz     | bmi     | -     | -     | 0
height  | ad      | 0   | 0       | -
height | aam | 0 | 0 | +
bmi     | height  | +     | 0     | 0
bmi     | cad     | +   | 0       | 0
bmi     | aam     | 0    | 0      | -
fn      | t2d     | 0     | 0     | + 
ls      | height  | +     | 0     | 0
aam | bmi | - | 0 | -
aam     | hdl     | 0     | 0     | +
aam     | ls      | 0       | 0   | -
aam     | height  | 0       | 0   | +
aam     | cd      | 0      | 0    | +
cd      | ad      | 0      | 0    | -
fg      | t2d     | + | 0 | +
tc      | bmi     | -   | 0       | 0
tc      | hdl     | -     | 0     | 0
tc      | cad     | + | 0 | +
tc      | ldl     | 0   | +        | +
hdl     | cad     | -   | -       | 0
ldl     | bmi     | - | 0 | -
ldl     | hdl     | -   | 0       | 0
ldl     | tg      | 0    | 0      | +
tg      | height  | -    | 0      | 0
tg      | bmi     | -     | 0     | 0
t2d     | hdl     | -    | 0      | 0
t2d     | cad     | +     | 0     | 0
ad      | tc      | +    | 0      | 0
ad      | hdl     | -     | 0     | 0
ad      | tg      | +     | 0     | 0

## Results using all 1.7 million SNPs

```{r, echo=FALSE}
res <- readRDS("../data/all_res.full_wts.RData")
pvals <- 2*pnorm(abs(res[1, , ]), lower.tail=FALSE)
logpvals <- -1*log10(pvals)

logpvals[(res[1, , ] < 0 & !is.na(res[1, , ])) ] <- -1 * logpvals[(!is.na(res[1, ,]) & res[1, ,] < 0)]

 dat.m <- melt(t(logpvals))
 dat.m$value[dat.m$value > 10] <- 10
 dat.m$value[dat.m$value < -10] <- -10
 dat.m$value[abs(dat.m$value) < 2] <- 0
 pbic <- ggplot(dat.m, aes(x=Var1, y=Var2, fill=value)) + 
   geom_tile() + coord_equal() + ggtitle("Not pruned, LDAK weighs, Best model(BIC)") + 
   scale_fill_gradient2(low="blue", high="red") + 
   theme(axis.text.x = element_text(angle = 90, hjust = 1))
 
 
 pvals <- 2*pnorm(abs(res[2, , ]), lower.tail=FALSE)
logpvals <- -1*log10(pvals)

logpvals[(res[2, , ] < 0 & !is.na(res[2, , ])) ] <- -1 * logpvals[(!is.na(res[2, ,]) & res[2, ,] < 0)]

 dat.m <- melt(t(logpvals))
 dat.m$value[dat.m$value > 10] <- 10
 dat.m$value[dat.m$value < -10] <- -10
 dat.m$value[abs(dat.m$value) < 2] <- 0
 plrt <- ggplot(dat.m, aes(x=Var1, y=Var2, fill=value)) + 
   geom_tile() + coord_equal() + ggtitle("Not Pruned, LDAK weighs, Best model(LRT)") + 
   scale_fill_gradient2(low="blue", high="red") + 
   theme(axis.text.x = element_text(angle = 90, hjust = 1))
 
 
 
 pvals <- 2*pnorm(abs(res[3, , ]), lower.tail=FALSE)
logpvals <- -1*log10(pvals)

logpvals[(res[3, , ] < 0 & !is.na(res[3, , ])) ] <- -1 * logpvals[(!is.na(res[3, ,]) & res[3, ,] < 0)]

 dat.m <- melt(t(logpvals))
 dat.m$value[dat.m$value > 10] <- 10
 dat.m$value[dat.m$value < -10] <- -10
 dat.m$value[abs(dat.m$value) < 2] <- 0
 pfull <- ggplot(dat.m, aes(x=Var1, y=Var2, fill=value)) + 
   geom_tile() + coord_equal() + ggtitle("Not Pruned, LDAK weighs, Full model") + 
   scale_fill_gradient2(low="blue", high="red") + 
   theme(axis.text.x = element_text(angle = 90, hjust = 1))
 
 grid.arrange(pbic, plrt, pfull, ncol=2)
```


