---
title: "Analysis of 16 GWAS traits"
author: "Jean Morrison"
date: "May 18, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

We analyzed 16 of the GWAS traits analyzed by Pickerell et al. 

## Summary of results 

```{r, echo=FALSE}
library(ggplot2)
library(reshape2)
res <- readRDS("../data/all_res.prune_wts.RData")
pvals <- 2*pnorm(abs(res[2, , ]), lower.tail=FALSE)
logpvals <- -1*log10(pvals)

logpvals[(res[1, , ] < 0 & !is.na(res[1, , ])) ] <- -1 * logpvals[(!is.na(res[1, ,]) & res[1, ,] < 0)]
logpvals[!is.na(pvals) & res[2,,] > 30] <- max(logpvals[!is.na(pvals) & res[2,,] < 30], na.rm=TRUE)

 dat.m <- melt(t(logpvals))
 dat.m$value[dat.m$value > 10] <- 10
 dat.m$value[dat.m$value < -10] <- -10
 dat.m$value[abs(dat.m$value) < 2] <- 0
 ggplot(dat.m, aes(x=Var1, y=Var2, fill=value)) + 
   geom_tile() + coord_equal() + ggtitle("Pruned") + 
   scale_fill_gradient2(low="blue", high="red") + 
   theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


```{r, echo=FALSE}
library(ggplot2)
library(reshape2)
res <- readRDS("../data/all_res.full_wts.RData")
pvals <- 2*pnorm(abs(res[2, , ]), lower.tail=FALSE)
logpvals <- -1*log10(pvals)

logpvals[(res[1, , ] < 0 & !is.na(res[1, , ])) ] <- -1 * logpvals[(!is.na(res[1, ,]) & res[1, ,] < 0)]
logpvals[!is.na(pvals) & res[2,,] > 30] <- max(logpvals[!is.na(pvals) & res[2,,] < 30], na.rm=TRUE)

 dat.m <- melt(t(logpvals))
 dat.m$value[dat.m$value > 10] <- 10
 dat.m$value[dat.m$value < -10] <- -10
 dat.m$value[abs(dat.m$value) < 2] <- 0
 ggplot(dat.m, aes(x=Var1, y=Var2, fill=value)) + 
   geom_tile() + coord_equal() + ggtitle("All") + 
   scale_fill_gradient2(low="blue", high="red") + 
   theme(axis.text.x = element_text(angle = 90, hjust = 1))
```



An incomplete list of differences from Pickerell et al:

Trait 1 | Trait 2 | $\lambda$ (stat) | Pickerell et al
--------|---------|------------------|-----------------
scz     | bmi     | -0.027 (-6.56)   | 0
scz     | hdl     | -0.017 (-2.17)   | 0
height  | bmi     | -0.054 (-9.26)   | 0
height  | cd      | 0.115 (3.10)     | 0
hieght  | fn      | 0.019 (1.57)     | 0
bmi     | cad     | 0.36 (4.62)      | 0
bmi     | tc      | -0.11 (-3.72)    | 0
bmi     | fg      | 0.052 (2.61)     | 0
bmi     | height  | 0.15 (4.97)      | 0 
fn      | height  | 0.071 (2.98)     | 0
fn      | bmi     | 0.044(2.83)      | 0
fn      | ad      | 0.20 (2.92)      | 0 
ls      | ad      | +              | 0
ls      | height  | +              | 0
aam     | cad     | -0.056 (-3.98)   | 0
aam     | height  | 0         | + 
aam     | cd   | 0           | + 
aam     | ls  | 0            | -
tg      | height  | -              | 0
tg      | bmi     | -              | 0
tg      | t2d     | pos              | 0 
ad      | bmi,tc,hdl,ldl,tg |         |0
t2d     |hdl,cad,cd |            | 0
cad     | cd, fg  |  +, +            | 0 
tg      | height, bmi, t2d| -,-,+ | 0
ldl     | hdl     | +  | 0
hdl     | aam     | 0 | +
hdl     | ldl, fg, cad | | 0
tc      | height, bmi, hdl| -, -, - | 0
fg | height | + | 0
cd | fn | - | 0
