---
title: "MAGMA"
author: "Nicholas Knoblauch"
date: "2018-12-13"
output: html_document
---


## MAGMA Results

Below is a summary of the output of MAGMA run on the GWAS summary statistics from `egg_bw_summary_statistics.tsv.gz` (with the sex-chromosome SNPs removed). MAGMA was run using the online `FUMA` server, specifically the `SNP2GENE` service ([link here](http://fuma.ctglab.nl))  The results can be found here: http://fuma.ctglab.nl/browse/48 (some of the plots are quite dense and might slow down your browser).

The table below is (by default) sorted by `prob_Z1_conf`.  The gene assignment was performed by `MAGMA`, and the p-value indicates the (bonferroni-adjusted) significance of the enrichment of the pathway in the birthweight GWAS. 

```{r,echo=F,message=F,warning=FALSE}

#First load libraries and 
library(tidyverse)
library(fs)

snp_annovf <- "~/Downloads/data/FUMA/annov.txt"
ind_snpf <- "~/Downloads/data/FUMA/IndSigSNPs.txt"
cause_file <- "/project/compbio/cause_results/gwas_pairs2/egg_bw__diagram_t2d_learn_cause.RDS"
full_gwas_f <- "/project/mstephens/data/external_public_supp/gwas_summary_statistics/summary_statistics/egg_bw_summary_statistics.tsv.gz"
```





```{r,echo=F,message=F,warning=FALSE}
#read in annovar results
annov_df <- read_tsv(snp_annovf) 
#read in GWAS input (to get rsid from annov
annov_df <- read_tsv(full_gwas_f,
                        col_types = cols_only(chrom="c",pos="i",snp="c",ref_allele="c",alt_allele="c")) %>%    mutate(uniqID = glue::glue("{chrom}:{pos}:{ref_allele}:{alt_allele}")) %>% select(uniqID,snp) %>% inner_join(annov_df)


#read in MAGMA significant SNPs
snp_df <- read_tsv("~/Downloads/data/FUMA/snps.txt")
annov_df <- select(snp_df,uniqID,rsID) %>% inner_join(annov_df)
#read in MAGMA genes
gene_df <- read_tsv("~/Downloads/data/FUMA/genes.txt")
#read in cause results
cause_data <- readRDS(cause_file)
cause_df <- cause_data[["data"]] %>% arrange(desc(prob_Z1_conf)) 
median_Z <- median(cause_df$prob_Z1_conf)
n_z <- sum(cause_df$prob_Z1_conf>median_Z)
```

```{r,echo=F,message=F,warning=FALSE}
cause_annov_df <- select(annov_df,snp=rsID,gene,symbol) %>% inner_join(select(cause_df,-seb1,-seb2,-A1,-A2,-delta_elpd)) %>% arrange(desc(prob_Z1_conf))
new_nz <- sum(cause_annov_df$prob_Z1_conf>median_Z)
```


## 
```{r}

```


```{r,echo=F,message=F,warning=FALSE}

magma_sets <- scan("~/Downloads/data/FUMA/magma.setgenes.out",what=character(),sep="\n")

magma_headers <- magma_sets[grepl("#",magma_sets)]

magma_data <- magma_sets[!grepl("#",magma_sets)] %>% str_split_fixed(pattern=regex("\\h+"),n = 10)
magma_cn <- c(magma_data[1,])
magma_data <- magma_data[magma_data[,2]!="GENE",]
magma_df <- as_data_frame(magma_data) %>% magrittr::set_colnames(magma_cn) %>% rename(SET=`_SET1_`) %>% select(SET,GENE,NSNPS,NPARAM,N,ZSTAT,P) %>%  mutate_at(c("ZSTAT","P","NSNPS","NPARAM","N"),as.numeric) %>% mutate(SET=as.integer(gsub("_SET([0-9]+)_","\\1",SET))) %>% rename(P_GENE=P,gene=GENE)

magma_header_mat <- t(matrix(magma_headers[-(1:2)],nrow=3))
magma_header_df <- gsub(".+= ","",magma_header_mat) %>% as_data_frame %>% magrittr::set_colnames(c("setname","ngenes","P")) %>% mutate(ngenes=as.integer(ngenes),P=as.numeric(P)) %>% mutate(SET=1:n()) %>% rename(P_SET=P)

magma_df <- inner_join(magma_df,magma_header_df) %>% arrange(P_GENE)
```




```{r,echo=F,message=F,warning=FALSE}
snp_cause_df <- mutate(gene_df,IndSigSNP=str_split(IndSigSNPs,pattern=";")) %>% 
  unnest() %>%
  select(gene=ensg,symbol,HUGO,IndSigSNP) %>% 
  inner_join(select(snp_df,IndSigSNP,snp=rsID),by=c("IndSigSNP")) %>% 
  select(-IndSigSNP) %>% 
  inner_join(select(cause_df,-seb1,-seb2,-A1,-A2,-delta_elpd)) %>%  
  inner_join(magma_df) %>%
  select(-ngenes,-P_GENE,-ZSTAT,-NSNPS,-SET,-NPARAM,-N) %>% 
  arrange(desc(prob_Z1_conf))
```
```{r,echo=F,message=F,warning=FALSE}
snp_cause_df<- distinct(snp_cause_df,setname,P_SET) %>% 
  mutate(set_adj_p=p.adjust(P_SET,method ="bonferroni",n= 10894)) %>% 
  select(-P_SET) %>% 
  inner_join(snp_cause_df)

select(snp_cause_df,
       gene=symbol,
       snp,beta_hat_BW=beta_hat_1,
       gene_set_adj_p=set_adj_p,
       beta_hat_t2d=beta_hat_2,
       prob_Z1_conf,
       gene_set=setname) %>% 
  DT::datatable()
```

```{r,echo=F,message=F,warning=FALSE}
median_Z <- median(cause_df$prob_Z1_conf)
# label snps with P(Z=1)>median_Z as "shared""
sub_cause_df <- mutate(cause_df,is_shared=prob_Z1_conf>median_Z) %>% 
  select(rsID=snp,is_shared,prob_Z1_conf)

#Merge the CAUSE results with the rest of the annotated GWAS significant results
magma_snp_df <- inner_join(annov_df,magma_df,by="gene") %>% 
  left_join(sub_cause_df,by="rsID") %>% #left join will assign NA to is_shared for gwas results that CAUSE didn't look at 
  mutate(is_shared=if_else(is.na(is_shared),FALSE,is_shared))

# What fraction of GWAS results are shared?
prob_shared <- distinct(magma_snp_df,rsID,is_shared) %>% 
  summarise(prob_shared=sum(is_shared)/n()) %>%
  pull(1)
```



##Enrichment

One way to frame the question of enrichment would be to ask: "Given that the proportion of Birthweight GWAS significant SNPs with $P(Z=1) > \text{Median}(P(Z=1))$ is `r prob_shared`, are there birthweight enriched pathways that have more SNPs with $P(Z=1) > \text{Median}(P(Z=1))$ SNPs than we would expect by chance alone?". Under the null hypothesis (or at least _a_ null hypothesis), the number of SNPs with $P(Z=1) > \text{Median}(P(Z=1))$ for a particular pathway is binomially distributed with the number of trials equal to the number of Birthweight GWAS significant SNPs in genes that are members of that pathway, and the probability of sucess is simply the proportion of _all_ GWAS SNPs for which $P(Z=1) > \text{Median}(P(Z=1))$. One simple test of enrichment would be a one-sided binomial test per pathway, followed by bonferroni correction.




```{r}
#Calculate the median P(Z=1)
median_Z <- median(cause_df$prob_Z1_conf)
# label snps with P(Z=1)>median_Z as "shared""
sub_cause_df <- mutate(cause_df,is_shared=prob_Z1_conf>median_Z) %>% 
  select(rsID=snp,is_shared,prob_Z1_conf)

#Merge the CAUSE results with the rest of the annotated GWAS significant results
magma_snp_df <- inner_join(annov_df,magma_df,by="gene") %>% 
  left_join(sub_cause_df,by="rsID") %>% #left join will assign NA to is_shared for gwas results that CAUSE didn't look at 
  mutate(is_shared=if_else(is.na(is_shared),FALSE,is_shared)) #Assign FALSE to the GWAS results that CAUSE didn't look at 



# For each gene set, count the total number of SNPs, and the number of SNPs above the median
set_sum <- group_by(magma_snp_df,setname) %>%
  summarise(n_snp_total=n_distinct(rsID), #total number of SNPs per pathway
            total_shared=sum(is_shared)) #shared SNPs per pathway

set_sum <- mutate(set_sum,
                  pval=pbinom(total_shared,size=n_snp_total,prob=prob_shared,lower.tail=F),
                  adj_p=p.adjust(pval,method="bonferroni")) %>% 
  arrange(adj_p) %>% rename(Pathway=setname,num_snps=n_snp_total)

set_sum %>%  DT::datatable()

```


