---
title: "Comparison of different priors in b-q model"
author: "Jean Morrison"
date: "October 9, 2017"
output:
  html_document: default
  pdf_document: default
---

## Introduction

In this document I will compare different priors for fitting the b-q model described [here](mh_test.html). Here we are primarily focused on fitting the shared model.

For all experiments, we place a prior on $z = tanh^{-1}(\rho)$ of $z \sim N(0, 0.5)$ and the MH sampler will sample values of $z$ rather than $\rho$. We will alsways run three chains of the sampler with length 4000 samples. We discard the first 2000 samples as burn-in and thin the remaining samples to retain only every 5th sample. The grids of mixture proportions were pre-computed as the MAP estimates.


For these experiments I will use five simulated data sets. Each is simulated under the b-q model with b=0.5 and $q$ equal t0 $0, 0.25, 0.5, 0.75$ and 1. These are single realizations from an early set of simulations I ran described [here](mh_test.html#experiments). These simulations have fairly strong signal and have on average 500 variants effecting trait 1 and 1000 variants effecting trait 2. 

```{r, eval=FALSE, echo=FALSE}
library(ashr)
library(numDeriv)
res <- list()
q <- c(0, 0.25, 0.5, 0.75, 1)
z_prior_func <- function(z){ dnorm(z, 0, 0.5, log=TRUE)}
for(i in 1:5){
  D <- readRDS(paste0("../data/test_data_easy/q", q[i], "_draw.RDS"))
  dat <- D$dat
  
  b1.fit <- ash(betahat = dat$beta_hat_1, sebetahat = dat$seb1,
              mixcompdist = "normal", prior="nullbiased")

  b2.fit <- ash(betahat = dat$beta_hat_2, sebetahat = dat$seb2,
              mixcompdist = "normal", prior="nullbiased")
  sigma1 <- b1.fit$fitted_g$sd[!zapsmall(b1.fit$fitted_g$pi)==0]
  sigma2 <- b2.fit$fitted_g$sd[!zapsmall(b2.fit$fitted_g$pi)==0]
  G <- data.frame("S1"=rep(sigma1, length(sigma2)),
                "S2"=rep(sigma2, each=length(sigma1)),
                "pi"=rep(0, length(sigma1)*length(sigma2)))
  arctanh <- function(rho){
    0.5*log((1+rho)/(1-rho))
  }
  map_null <- get_map(pars.start=c(0),
                    data=dat, grid=G,
                    par.bounds=cbind(c(-1), c(1)),
                    type="null", z_prior_func= z_prior_func, n.iter=20)
  grid <- map_null$grid
  grid$pi <- map_null$pi
  saveRDS(grid, file=paste0("../data/test_data_easy/q", q[i], "_grid.RDS"))
}
```


## Normal b, Beta q
Priors:

+ $b \sim N(0, 0.6)$
+ $q \sim Beta(0.1, 0.9)$

```{r, cache=TRUE, eval=FALSE, results="hide", one}
library(sherlockAsh)
#Set priors
z_prior_func <- function(z){ dnorm(z, 0, 0.5, log=TRUE)}
zsamp <- function(n){rnorm(n, 0, 0.5)}
b_prior_func <- function(b){dnorm(b, 0, 0.6, log=TRUE)}
bsamp <- function(n){rnorm(n, 0, 0.6)}
q_prior_func <- function(q){dbeta(q, 0.1, 0.9, log=TRUE)}
qsamp <- function(n){runif(n, 0, 1)}

#Starting variance of sampling distribution
prop_sigma <- c(1e-4, 1e-4, 1e-4)

res <- list()
q <- c(0, 0.25, 0.5, 0.75, 1)
for(i in 1:5){
  D <- readRDS(paste0("../data/test_data_easy/q", q[i], "_draw.RDS"))
  dat <- D$dat
  grid <- readRDS(paste0("../data/test_data_easy/q", q[i], "_grid.RDS"))
  res[[i]] <- mh_compare2(dat, grid, logitq=F, seed=1e8, models=3,
                   z_prior_func = z_prior_func, zsamp=zsamp,
                   b_prior_func = b_prior_func, bsamp=bsamp,
                   q_prior_func = q_prior_func, qsamp=qsamp,
                   prop_sigma = prop_sigma,
                   nchain=3, niter_perchain=4000, burn_in = 2000, thin=5,
                   return_mh=TRUE)
}
saveRDS(res, file="../data/test_data_easy/res1.RDS")
```

Trace plots for all five data sets

```{r, echo=FALSE, two}
library(sherlockAsh)
library(ggplot2)
library(tidyr)
res1 <- readRDS("../data/test_data_easy/res1.RDS")
q <- c(0, 0.25, 0.5, 0.75, 1)
n <- dim(res1[[1]]$traces)[2]/3
tr_1 <- cbind(res1[[1]]$traces[3,,], rep(1:3, each=n), rep(1:n, 3), rep(q[1], 3*n))
for(i in 2:5){
  tr1 <- cbind(res1[[i]]$traces[3,,], rep(1:3, each=n), rep(1:n, 3), rep(q[i], 3*n))
  tr_1 <- rbind(tr_1, tr1)
}
tr_1 <- data.frame(tr_1)
names(tr_1) <- c("z", "b", "q", "chain", "iter", "trueq")
tr_1 <- gather(tr_1, "variable", "value", -chain, -iter, -trueq)
tr_1$chain <- as.factor(tr_1$chain)
p <- ggplot(tr_1) + geom_line(aes(x=iter, y=value, group=chain, col=chain)) + 
  facet_wrap(~trueq + variable, scales="free", nrow=5, ncol=3) + theme_bw()
p
```

The following table shows 90\% credible intervals for each of the three parameters as well as acceptance rates for all three chains.
```{r, echo=FALSE, three}
library(knitr)
ar <- sapply(res1, function(x){
  paste0(x$acceptance_rat[,3], collapse=",")
})
qci <- sapply(res1, function(x){
  yy <- as.numeric(quantile(x$traces[3,,3], probs = c(0.05, 0.95)))
  paste0(round(yy, digits=2), collapse=",")
})
bci <- sapply(res1, function(x){
  yy <- as.numeric(quantile(x$traces[3,,2], probs = c(0.05, 0.95)))
  paste0(round(yy, digits=2), collapse=",")
})
zci <- sapply(res1, function(x){
  yy <- as.numeric(quantile(x$traces[3,,1], probs = c(0.05, 0.95)))
  paste0(round(yy, digits=2), collapse=",")
})
df <- data.frame("q"=q, "acceptance_rate" = ar, "qci"=qci, 
                 "bci"=bci, "zci"=zci)
kable(df, caption="Normal b, Beta q")
```

This combination of priors can sometimes lead to low acceptance rates as in the case of the third chain with $q=0.75$. For the most part, the credible intervals overlap the target parameter values or are close. The credible intervals for $b$ are all slightly low (in this case $b = 0.5$) when $q > 0$.




## "Non-local" $b$, Normal $logit(q)$ sample from $logit(q)$
Priors:

+ $b$ non-local
$$
\pi_b(b) = \frac{b^2}{0.4^2}N(b; 0, 0.4)
$$
+ $logit(q) \sim N(0, 1.2)$

```{r, cache=TRUE, results="hide", eval=FALSE, four}
#Set priors
z_prior_func <- function(z){ dnorm(z, 0, 0.5, log=TRUE)}
zsamp <- function(n){rnorm(n, 0, 0.5)}
b_prior_func <- function(b){
  log((b/0.4)^2) + dnorm(b, 0, 0.4, log=TRUE)
}
bsamp <- function(n){rnorm(n, 0, 0.6)}
q_prior_func <- function(logitq){dnorm(logitq, 0, 1.2, log=TRUE)}
qsamp <- function(n){logit(runif(n, 0, 1))}

#Starting variance of sampling distribution
prop_sigma <- c(1e-4, 1e-4, 1e-4)

res <- list()
q <- c(0, 0.25, 0.5, 0.75, 1)
for(i in 1:5){
  D <- readRDS(paste0("../data/test_data_easy/q", q[i], "_draw.RDS"))
  dat <- D$dat
  grid <- readRDS(paste0("../data/test_data_easy/q", q[i], "_grid.RDS"))
  res[[i]] <- mh_compare2(dat, grid, logitq=T, seed=1e8, models=3,
                   z_prior_func = z_prior_func, zsamp=zsamp,
                   b_prior_func = b_prior_func, bsamp=bsamp,
                   q_prior_func = q_prior_func, qsamp=qsamp,
                   prop_sigma = prop_sigma,
                   nchain=3, niter_perchain=4000, burn_in = 2000, thin=5,
                   return_mh=TRUE)
}
saveRDS(res, file="../data/test_data_easy/res2.RDS")
```

Trace plots for all five data sets

```{r, echo=FALSE, five}
res2 <- readRDS("../data/test_data_easy/res2.RDS")
n <- dim(res2[[1]]$traces)[2]/3
tr_2 <- cbind(res2[[1]]$traces[3,,], rep(1:3, each=n), rep(1:n, 3), rep(q[1], 3*n))
for(i in 2:5){
  tr1 <- cbind(res2[[i]]$traces[3,,], rep(1:3, each=n), rep(1:n, 3), rep(q[i], 3*n))
  tr_2 <- rbind(tr_2, tr1)
}
tr_2 <- data.frame(tr_2)
names(tr_2) <- c("z", "b", "q", "chain", "iter", "trueq")
tr_2 <- gather(tr_2, "variable", "value", -chain, -iter, -trueq)
tr_2$chain <- as.factor(tr_2$chain)
p <- ggplot(tr_2) + geom_line(aes(x=iter, y=value, group=chain, col=chain)) + 
  facet_wrap(~trueq + variable, scales="free", nrow=5, ncol=3) + theme_bw()
p
```


```{r, echo=FALSE, six}
library(knitr)
ar <- sapply(res2, function(x){
  paste0(x$acceptance_rat[,3], collapse=",")
})
qci <- sapply(res2, function(x){
  yy <- as.numeric(quantile(x$traces[3,,3], probs = c(0.05, 0.95)))
  paste0(round(yy, digits=2), collapse=",")
})
bci <- sapply(res2, function(x){
  yy <- as.numeric(quantile(x$traces[3,,2], probs = c(0.05, 0.95)))
  paste0(round(yy, digits=2), collapse=",")
})
zci <- sapply(res2, function(x){
  yy <- as.numeric(quantile(x$traces[3,,1], probs = c(0.05, 0.95)))
  paste0(round(yy, digits=2), collapse=",")
})
df <- data.frame("q"=q, "acceptance_rate" = ar, "qci"=qci, 
                 "bci"=bci, "zci"=zci)
kable(df, caption="Non-local b, normal logit(q), sample logit(q)")
```

The credible intervals using this set of priors is quite similar to the ones in the previous section. They are a bit smaller and the sampler did not have as much trouble with low acceptance rates. Note that in this scheme, the intervals for $q$ cannot include 0 or 1 because we are sampling on the $logit(q)$ scale but have transformed back to the original scale to compute intervals.


## Normal b, Beta q, sampled on $logit(q)$ scale
I tried sampling from the posterior of $logit(q)$ using the first set of priors to see if this improves the acceptance rate, however, this had the opposite effect. I found that the variance of the proposal distribution tended to end up too large in the adaptive MH algorithm, leading to acceptance rates of zero:

```{r, cache=TRUE, results="hide", echo=FALSE, eval=FALSE, seven}
library(sherlockAsh)
#Set priors
z_prior_func <- function(z){ dnorm(z, 0, 0.5, log=TRUE)}
zsamp <- function(n){rnorm(n, 0, 0.5)}
b_prior_func <- function(b){dnorm(b, 0, 0.6, log=TRUE)}
bsamp <- function(n){rnorm(n, 0, 0.6)}
q_prior_func <- function(logitq){dbeta(expit(logitq), 0.1, 0.9, log=TRUE)}
qsamp <- function(n){logit(runif(n, 0, 1))}

#Starting variance of sampling distribution
prop_sigma <- c(1e-4, 1e-4, 1e-4)

res <- list()
q <- c(0, 0.25, 0.5, 0.75, 1)
for(i in 1:5){
  D <- readRDS(paste0("../data/test_data_easy/q", q[i], "_draw.RDS"))
  dat <- D$dat
  grid <- readRDS(paste0("../data/test_data_easy/q", q[i], "_grid.RDS"))
  res[[i]] <- mh_compare2(dat, grid, logitq=T, seed=1e8, models=3,
                   z_prior_func = z_prior_func, zsamp=zsamp,
                   b_prior_func = b_prior_func, bsamp=bsamp,
                   q_prior_func = q_prior_func, qsamp=qsamp,
                   prop_sigma = prop_sigma,
                   nchain=3, niter_perchain=4000, burn_in = 2000, thin=5,
                   return_mh=TRUE)
}
saveRDS(res, file="../data/test_data_easy/res3.RDS")
```


```{r, echo=FALSE, eval=FALSE, eight}
res3 <- readRDS("../data/test_data_easy/res3.RDS")
q <- c(0, 0.25, 0.5, 0.75, 1)
n <- dim(res3[[1]]$traces)[2]/3
tr <- cbind(res3[[1]]$traces[3,,], rep(1:3, each=n), rep(1:n, 3), rep(q[1], 3*n))
for(i in 2:5){
  tr1 <- cbind(res3[[i]]$traces[3,,], rep(1:3, each=n), rep(1:n, 3), rep(q[i], 3*n))
  tr <- rbind(tr, tr1)
}
tr <- data.frame(tr)
names(tr) <- c("z", "b", "q", "chain", "iter", "trueq")
tr <- gather(tr, "variable", "value", -chain, -iter, -trueq)
tr$chain <- as.factor(tr$chain)
p <- ggplot(tr) + geom_line(aes(x=iter, y=value, group=chain, col=chain)) + 
  facet_wrap(~trueq + variable, scales="free", nrow=5, ncol=3) + theme_bw()
p
```

```{r, echo=FALSE, nine}
res3 <- readRDS("../data/test_data_easy/res3.RDS")
q <- c(0, 0.25, 0.5, 0.75, 1)
ar <- sapply(res3, function(x){
  paste0(x$acceptance_rat[,3], collapse=",")
})
qci <- sapply(res3, function(x){
  yy <- as.numeric(quantile(x$traces[3,,3], probs = c(0.05, 0.95)))
  paste0(round(yy, digits=2), collapse=",")
})
bci <- sapply(res3, function(x){
  yy <- as.numeric(quantile(x$traces[3,,2], probs = c(0.05, 0.95)))
  paste0(round(yy, digits=2), collapse=",")
})
zci <- sapply(res3, function(x){
  yy <- as.numeric(quantile(x$traces[3,,1], probs = c(0.05, 0.95)))
  paste0(round(yy, digits=2), collapse=",")
})
df <- data.frame("q"=q, "acceptance_rate" = ar, "qci"=qci, 
                 "bci"=bci, "zci"=zci)
kable(df, caption="Normal b, Beta q, sampled on logit(q) scale")
```

## Comparing posterior and prior distributions of $q$

Plots below show histograms of posterior samples of $q$. The solid lines show the prior density.
```{r, echo=FALSE, ten}
q_prior_func1 <- function(q){dbeta(q, 0.1, 0.9)}

n <- with(tr_2, sum(variable=="q" & trueq==0))
bandwidth <- 1/30

q_prior_func2 <- function(q){dnorm(logit(q), 0, 1.2)*(1/(q-q^2))*n*bandwidth}
q_prior_func1 <- function(q){dbeta(q, 0.1, 0.9)*n*bandwidth}

ggplot(tr_1[tr_1$variable=="q",]) +
  geom_histogram(aes(x=value)) + facet_wrap("trueq") + 
  stat_function(fun=q_prior_func1, xlim=c(0, 1)) + theme_bw()

ggplot(tr_2[tr_2$variable=="q",]) +
  geom_histogram(aes(x=value)) + facet_wrap("trueq") + 
  stat_function(fun=q_prior_func2, xlim=c(0, 1)) + theme_bw()


```
