---
title: "More LD Simulations"
author: "Jean Morrison"
date: 2017-12-12
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->
## Questions to answer

+ In our previous simulations with LD, we found that we had very reduced power when we used a set of LD pruned SNPs in which we preferenctially chose SNPs with low trait 1 $p$-values. Why is this power low?
+ Can we improve power?
+ If we use a subset of SNPs, what is the best grid prior to use for the direct effects?

## Set up

I decided to use a simpler LD structure to answer these questions. I pulled one 10 SNP wide LD block out of 1,000 genomes. 
```{r,  echo=FALSE, message=FALSE, pl1}
library(reshape2)
library(ggplot2)
library(matrixStats)
library(tidyr)
library(dplyr)
ld <- readRDS("../data/small_ld.RDS")
melted_R <- melt(ld$c)
ggplot(melted_R) + geom_tile(aes(x=Var1, y=Var2, fill=value)) + scale_fill_gradient2(low="blue", high="red", mid="white")
```

Each simulated data set contains 10,000 blocks of 10 SNPs. Each block has the LD pattern shown above. In each block there can be at most one effect SNP.

## Analysis using all SNPs and oracle SNPs

First I compare two methods that don't rely on pruning SNPs:

+ All-nowts: Use all SNPs, calculate the grid prior using all SNPs
+ oracle block: Use one SNP in each block. In blocks containing trait 1 effect SNPs, use the true trait 1 effect SNP. Grid parameters using this same set of SNPs. 

The oracle block method shows the power we would expect in a similar problem with no LD. Ignoring LD in the "All-nowts" method results in somewhate higher type 1 error for some values of $q$.
```{r,echo=FALSE, plot1}
res_wide <- readRDS("../data/res_wide_2017-12-18.RDS")

pos_curves <- function(res_wide, methods, level=0.95){
    pos_tab <- data.frame("q"=unique(res_wide$q))
    for(m in methods){
        pos <- sapply(pos_tab$q, function(x){ with(res_wide, sum(get(m)[q==x] < -1*qnorm(level)))})
        pos_tab[,m] <- pos
    }
    pos_tab <- gather(pos_tab, "method", "num_pos", -q)
    pos_tab[,"level"] <- level
    return(pos_tab)
}
methods <- c("oracle block", "All-nowts")
pos_tab <- rbind(pos_curves(res_wide, methods, 0.95),
                  pos_curves(res_wide, methods, 0.90),
                  pos_curves(res_wide, methods, 0.975))
pos_tab$level <- factor(pos_tab$level)
plt <- ggplot(pos_tab) +
       geom_point(aes(x=q, y=num_pos, color=method, shape=level), size=1.3) +
       geom_line(aes(x=q, y=num_pos, color=method, linetype=method, group=paste0(level, method))) +
       ylab("Num. with sig. z-score in favor of causal model") +
       facet_wrap(~level) +
       theme_bw()
plt

```

## Best SNP per block

Now we consider

+ best block: Use one SNP per block. Choose the SNP with the lowest trait 1 $p$-value. Grid parameters calculated using all SNPs. 
+ best block/oracle: Use one SNP per block but only include blocks with true trait 1 effects. As above, choose the SNP with lowest trait 1 $p$-value. Same grid as above. 
+ oracle/oracle: Use only the true trait 1 effect SNPs. Grid from the set of all SNPs used in "oracle block"

```{r,echo=FALSE, plot2}
methods <- c("oracle block", "best block", "best block/oracle", "oracle/oracle")
pos_tab <- rbind(pos_curves(res_wide, methods, 0.95),
                  pos_curves(res_wide, methods, 0.90),
                  pos_curves(res_wide, methods, 0.975))
pos_tab$level <- factor(pos_tab$level)
plt <- ggplot(pos_tab) +
       geom_point(aes(x=q, y=num_pos, color=method, shape=level), size=1.3) +
       geom_line(aes(x=q, y=num_pos, color=method, linetype=method, group=paste0(level, method))) +
       ylab("Num. with sig. z-score in favor of causal model") +
       facet_wrap(~level) +
       theme_bw()
plt

```

The "oracle/oracle" method should be the best. It includes only trait 1 effect SNPs and has a good set of SNPs to estimate the grid with. There shouldn't be bias introduced by selection. This method is the most powerful but also has more type 1 errors. Disturbingly, we have more type 1 errors at $q=0$. 

"best block/oracle" is very similar to the "oracle/oracle" method. This shows that the loss of power can't be explained by choosing the wrong SNP in effect blocks. The "best block" method has lower power. This must be a result of including SNPs that aren't in trait 1 effect blocks but have higher trait 1 effect estimates as a result of selection bias. 

## More stringent selection

Based on these results, I experimented with using a more stringent threshold and including fewer SNPs when calculating the posterior distribution. 

+ best block/top: Use the best trait 1 SNP in each block but keep only the top 100 blocks. Grid calculated from all SNPs. 

```{r,echo=FALSE, plot3}
methods <- c("oracle block", "best block", "best block/top")
pos_tab <- rbind(pos_curves(res_wide, methods, 0.95),
                  pos_curves(res_wide, methods, 0.90),
                  pos_curves(res_wide, methods, 0.975))
pos_tab$level <- factor(pos_tab$level)
plt <- ggplot(pos_tab) +
       geom_point(aes(x=q, y=num_pos, color=method, shape=level), size=1.3) +
       geom_line(aes(x=q, y=num_pos, color=method, linetype=method, group=paste0(level, method))) +
       ylab("Num. with sig. z-score in favor of causal model") +
       facet_wrap(~level) +
       theme_bw()
plt

```

This substantially improved power but it also increased the false positive rate. 

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
