---
title: "Co-regulation simulations"
author: "Jean Morrison"
date: "May 15, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

These simulations are intended to understand the effects of co-regulation on our method. The simulation set-up is as follows:

There is some un-known, un-measured factor $F$ that influences both trait 1 and trait 2. Additionally, trait 1 directly (or through a pathway separate from $F$) affects trait 2. 

## Data generation

I will vary two parameters:
+ $q$: The proportion of trait 1 effects that come through the shared factor
+ $b$: The effect size of the shared factor on trait 2

```{r, one}
library(ashr)
library(sherlockAsh)
library(tidyr)

q <- 0.2
b <- 0.13

p_f <- 0.07*q
p_1 <- 0.07*(1-q)

#Function for effects on factor
gfn <- normalmix(pi=c(1-p_f, p_f),
                mean=rep(0, 2),
                sd=c(0,  0.1))
gf <- function(p){ rnormalmix(p, gfn)}


g1n <- normalmix(pi=c(1-p_1, p_1),
                mean=rep(0, 2),
                sd=c(0, 0.1))
g2n <- normalmix(pi=c(0.9, 0.1),
                        mean=rep(0, 2),
                        sd=c(0, 0.05))

g1 <- function(bf){
    bf + rnormalmix(length(bf), g1n)
}
g2 <- function(bf, b1){
    b*bf + lambda *b1 +  rnormalmix(length(bf), g1n)
}

maf.func <- function(p){rbeta(n=p, 1, 5)}

```

## Results

```{r}
library(ggplot2)
qs <-  c(0.05, 0.1, 0.2, 0.4, 0.8, 1)
bs <- c(0.05, 0.13)

t1e <- matrix(nrow=length(qs), ncol=2)
for(i in 1:length(qs)){
  for(j in 1:2){
    res <- readRDS(paste0("../data/res.", qs[i], ".", bs[j], ".0.RData"))
    stats <- res$corr[3,,1]/res$corr[3,,2]
    t1e[i,j] <- mean(abs(stats) > abs(qnorm(0.975, lower.tail=TRUE)))
  }
}

t1e <- data.frame(t1e)
names(t1e) <- c("b_0.05", "b_0.13")
t1e$q <- qs
t1e_long <- gather(t1e, "b", "Type1Error", -q)

n <- 100

c975 <- qbinom(p=0.975, size=n, prob=0.05)/n

c025 <- qbinom(p=0.025, size=n, prob=0.05)/n
v <- seq(0, 1, length.out=1000)
df.shade = data.frame("x"=c(v, rev(v)), "y"=c(rep(c025, 1000), rep(c975, 1000)))


ggplot(t1e_long) +
  geom_polygon(data=df.shade, aes(x=x, y=y), alpha=0.1, fill="black") +
  geom_point(aes(x=q, y=Type1Error, group=b, color=b, shape=b), size=2) +
  geom_line(aes(x=q, y=Type1Error, group=b, color=b, lty=b)) +
  geom_hline(yintercept = 0.05, lty=3) +
  xlab("Proportion of Type 1 Effects\nFrom Shared Factor") + ylab("Type 1 Error") + ggtitle("Type 1 Error") +
  theme_bw()
```
